Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa

### 2022-10-06 src/docs_06-10-22-hdsl/06-10-22-hdsl.Rmd
Update HDSL talk

* In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are disproportionately at risk of HIV infection
* This disparity is because of:
  1. Younger age at first sex
  2. Age patterns of sexual mixing
  3. Structural vulnerabilities and power imbalances
  4. Increased susceptibility to HIV infection

@@

* We used sexual behaviour data from AIS, BAIS, DHS and PHIA household surveys to place respondents into $K = 4$ risk groups:
  1. $k = 1$ Not sexually active
  2. $k = 2$ One cohabiting sexual partner
  3. $k = 3$ Non-regular sexual partner(s)
  4. $k = 4$ Female sex workers
* District-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi]
* Risk ratios from ALPHA network analysis [@slaymaker2020croi] and UNAIDS analysis led by Keith Sabin

@@

* Would like to use integrated nested Laplace approximations for fast, accurate inference, but `R-INLA` is not compatible with multinomial likelihoods because they depend on multiple elements of the latent field
* Instead, use that multinomial logistic regression models can be recast as a Poisson log-linear models using the Poisson trick
* Include observation-specific random effects $\theta_{ita} \sim \mathcal{N}(0, 1000^2)$ in any linear predictor $\eta_{itak} = \theta_{ita} + \cdots$ to ensure exact reproduction of $m_{ita}^\star$

>>>>

* In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are disproportionately at risk of HIV infection
  * 28% of population and 44% of new infections
* This disparity is because of:
  1. Younger age at first sex
  2. Age patterns of sexual mixing
  3. Structural vulnerabilities and power imbalances
  4. Increased susceptibility to HIV infection

```{r echo=FALSE, out.width = "125px", fig.cap="Tweet from UNAIDS"}
knitr::include_graphics("tweet.png")
```

@@

* We used sexual behaviour data from AIS, BAIS, DHS and PHIA household surveys to place respondents into $K = 4$ risk groups:
  1. $k = 1$ Not sexually active
  2. $k = 2$ One cohabiting sexual partner
  3. $k = 3$ Non-regular sexual partner(s)
  4. $k = 4$ Female sex workers
* District-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi]
  * Combines survey and programmatic data to estimate indicators in the general population
* Risk ratios from ALPHA network analysis [@slaymaker2020croi] and UNAIDS analysis led by Keith Sabin
  * 10 longitudinal studies: https://alpha.lshtm.ac.uk/

@@

* Would like to use integrated nested Laplace approximations for fast, accurate inference, but `R-INLA` is not compatible with multinomial likelihoods because they depend on multiple elements of the latent field
* Instead, use that multinomial logistic regression models can be recast as a Poisson log-linear models using the Poisson trick
* This works because conditional on their sum, Poisson counts are jointly multinomially distributed
* If you include observation-specific random effects $\theta_{ita} \sim \mathcal{N}(0, 1000^2)$ in the linear predictor $\eta_{itak} = \theta_{ita} + \cdots$ then the sample sizes $m_{ita}^\star$ can be exactly recovered, ensuring the models are actually the same
